```{r,echo=FALSE, message = FALSE}

library(tidyverse)
library(gapminder)
library(stringr)

data = gapminder::gapminder


perm_mean <- function(perms = 1000, values, n1)
{
  ## Variables ##
  # perms: The number of permutations 
  # values (num): 
  # n1 (int): Size of group 1
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # Permuted mean difference 
  zerovector = numeric(660)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    test <- sample(values, length(values), replace = F)
    group1 <- test[1:n1]
    group2 <- test[1 + n1:length(values) -1 ]
    # Step 3:
    # Compute the sample means for the two groups from 
    Meangroup1 <- mean(group1)
    Meangroup2 <- mean(group2)
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    difference <- (Meangroup2 - Meangroup1)
    zerovector[i] = difference
  }
  
  # Step 5:
  # Return new updated vector, created in step 1
  return(zerovector)
}
```






#How similar are the life expectancies between Europe and Americas

```{r,echo=FALSE, message = FALSE}
AmericasEurope <- data %>%
  filter(continent == "Americas" | continent ==  "Europe") %>%
  select(-'pop',-'gdpPercap')

DataForPlot <- perm_mean(perms = 660, AmericasEurope$lifeExp,350)
view(DataForPlot)
Dataneeded <- cbind(AmericasEurope, DataForPlot)
ggplot(data = AmericasEurope) +
  geom_density(mapping = aes(x = DataForPlot, color = continent)) +
  geom_vline(xintercept = 0.01273659, color = 'blue') +
  geom_vline(xintercept = -.41924634, color = 'red') +
  geom_vline(xintercept = .42350648, color = 'purple') +
  geom_vline(xintercept = .5922599, color = 'black') +
  geom_vline(xintercept = -.5922599, color = 'black')
view(quantile(DataForPlot))

```



As we can see, the 50% is nearly zero for our permutation test between the two continents.
This tells us that the two continents life expectancies are indeed similar
Shown on the plots are the other values of the 25, and 75%.
The two densities are very similar.
The min value is a difference of -2.04, while the max is a difference of 1.94)
These values most likely represent countries of significant differences in quality of life.
Overall, we  conclude the life expectancies are indeed very similar. 
The black lines represent the pvalue.


#Calculating P value
```{r}
a <- 0
s <- sd(Dataneeded$DataForPlot)
n <- 660
xbar <- mean(Dataneeded$DataForPlot)
z <- (xbar - a)/(s/sqrt(n))

pvalue <- 2*pnorm(-abs(z))


```


Our p-value gives us a rather large probability of an observation assuming our null hypothesis is true, so we accept this.  Our p value has a value of .5922599